---
title: "Data Visualization by Kieran Healy"
author: "Michal Dabrowski"
date: "2021-01-27 - 2021-02-11"
output: html_document
---
   
## Data Visualization Notes

This is a starter RMarkdown template to accompany *Data Visualization* (Princeton University Press, 2019). You can use it to take notes, write your code, and produce a good-looking, reproducible document that records the work you have done. At the very top of the file is a section of *metadata*, or information about what the file is and what it does. The metadata is delimited by three dashes at the start and another three at the end. You should change the title, author, and date to the values that suit you. Keep the `output` line as it is for now, however. Each line in the metadata has a structure. First the *key* ("title", "author", etc), then a colon, and then the *value* associated with the key.  

## This is an RMarkdown File

Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. A *code chunk* is a specially delimited section of the file. You can add one by moving the cursor to a blank line choosing Code > Insert Chunk from the RStudio menu. When you do, an empty chunk will appear:

```{r}


```

Code chunks are delimited by three backticks (found to the left of the 1 key on US and UK keyboards) at the start and end. The opening backticks also have a pair of braces and the letter `r`, to indicate what language the chunk is written in. You write your code inside the code chunks. Write your notes and other material around them, as here. 

## Before you Begin

To install the tidyverse, make sure you have an Internet connection. Then *manually* run the code in the chunk below. If you knit the document if will be skipped. We do this because you only need to install these packages once, not every time you run this file. Either knit the chunk using the little green "play" arrow to the right of the chunk area, or copy and paste the text into the console window.

```{r install, eval = FALSE}

## This code will not be evaluated automatically.
## (Notice the eval = FALSE declaration in the options section of the
## code chunk)

my_packages <- c("tidyverse", "broom", "coefplot", "colorblindr" "cowplot",
                 "dichromat" "gapminder", "GGally", "ggrepel", "ggridges",
                 "ggthemes", "gridExtra", "here", "interplot", "margins",
                 "maps", "mapproj", "mapdata", "MASS", "quantreg",
                 "rlang", "scales", "survey", "srvyr", "statebins",
                 "viridis", "viridisLite", "devtools")

install.packages(my_packages, repos = "http://cran.rstudio.com")

devtools::install_github("kjhealy/socviz", "hrbrmstr/hrbrthemes", "kjhealy/myriad")

```

## Set Up Your Project and Load Libraries

To begin we must load some libraries we will be using. If we do not load them, R will not be able to find the functions contained in these libraries. The tidyverse includes ggplot and other tools. We also load the socviz and gapminder libraries.

```{r setup, include=FALSE}

## By default, show code for all chunks in the knitted document,
## as well as the output. To override for a particular chunk
## use echo = FALSE in its options.
knitr::opts_chunk$set(echo = TRUE) 

## Set the default size of figures
knitr::opts_chunk$set(fig.width = 8, fig.height = 5)  

## Load the libraries we will be using
library(gapminder)
library(here)
library(socviz)
library(tidyverse)

```

Notice that here, the braces at the start of the code chunk have some additional options set in them. There is the language, `r`, as before. This is required. Then there is the word `setup`, which is a label for your code chunk. Labels are useful to briefly say what the chunk does. Label names must be unique (no two chunks in the same document can have the same label) and cannot contain spaces. Then, after the comma, an option is set: `include=FALSE`. This tells R to run this code but not to include the output in the final document. 


When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r look}
gapminder
```

The remainder of this document contains the chapter headings for the book, and an empty code chunk in each section to get you started. Try knitting this document now by clicking the "Knit" button in the RStudio toolbar, or choosing File > Knit Document from the RStudio menu.

## Look at Data

```{r}

my_numbers <- c(1, 1, 4, 1, 1, 4, 1)
my_numbers
4 + 1
letters

```

## Get Started

```{r}

my_numbers <- c(1, 2, 3, 1, 3, 5, 25)
your_numbers <- c(5, 31, 71, 1, 3, 21, 6)

my_numbers

mean(x = my_numbers)
mean(x = your_numbers)

mean(my_numbers)

my_summary <- summary(my_numbers)
my_summary

table(my_numbers)

sd(my_numbers)
my_numbers*5
my_numbers + 1
my_numbers + your_numbers

class(my_numbers)
class(my_summary)
class(summary)

my_new_vector <- c(my_numbers, "Apple")
my_new_vector
class(my_new_vector)

titanic
class(titanic)
titanic$percent

titanic_tb <- as_tibble(titanic)
titanic_tb

str(my_numbers)
str(my_summary)

url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"
organs <- read_csv(file = url)

```

## Make a Plot

```{r}

library(gapminder)
gapminder

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_smooth()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + geom_smooth()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + geom_smooth(method = "lm")

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + geom_smooth(method = "gam") + scale_x_log10()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + geom_smooth(method = "lm") +
    scale_x_log10(labels = scales::dollar)

```
```{r}

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = "purple"))

p + geom_point() + 
    geom_smooth(method = "loess") +
    scale_x_log10()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(color = "purple") + 
    geom_smooth(method = "loess") + 
    scale_x_log10()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(alpha = 0.3) + 
    geom_smooth(color = "orange", se = FALSE, size = 2, method = "lm") +
    scale_x_log10()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(alpha = 0.3) + 
    geom_smooth(method = "gam") +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapiminder.")

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = continent))

p + geom_point() + 
    geom_smooth(method = "loess") + 
    scale_x_log10()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = continent, fill = continent))

p + geom_point() + 
    geom_smooth(method = "loess") +
    scale_x_log10()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))

p + geom_point(mapping = aes(color = continent)) + 
    geom_smooth(method = "loess") +
    scale_x_log10()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point(mapping = aes(color = log(pop))) + scale_x_log10()

```
```{r example, fig.width = 12, fig.height = 9}

p_out <- p + geom_point() + geom_smooth(method = "loess") + scale_x_log10()
ggsave(filename = "my_figure.png", plot = p_out)

here()
ggsave(here("figures", "life_vs_gdp_gradient.pdf"), plot = p_out, 
       height = 8, width = 10, units = "in")

```
```{r}

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + geom_smooth()
p + geom_smooth() + geom_point()

p <- ggplot(data = gapminder, mapping = aes(x = pop, y = lifeExp))
p + geom_point()
p + geom_point() + scale_x_log10()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp))
p + geom_point() + scale_x_sqrt()
p + geom_point() + scale_x_reverse()
p + geom_point() + scale_y_log10()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = year))
p + geom_point()

gapminder

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = factor(year)))
p + geom_point()

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = factor(year)))

p + geom_point() + 
    geom_smooth(se = FALSE, method = "lm") +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapiminder.")

```
```{r example2, fig.width = 10, fig.height = 12}

p <- ggplot(data = gapminder, mapping = aes(x = gdpPercap, y = lifeExp, color = country))

p + geom_point() + 
    geom_smooth(se = FALSE, method = "lm") +
    scale_x_log10(labels = scales::dollar) +
    labs(x = "GDP Per Capita", y = "Life Expectancy in Years",
         title = "Economic Growth and Life Expectancy",
         subtitle = "Data points are country-years",
         caption = "Source: Gapiminder.") +
    theme(legend.position = "bottom")

```

## Show the Right Numbers

```{r}

p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line()

p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line(aes(group = country))

p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p + geom_line(aes(group = country)) + facet_wrap(~ continent)

p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))

p + geom_line(color = "gray70", aes(group = country)) +
    geom_smooth(size = 1.1, method = "loess", se = FALSE) + 
    scale_y_log10(labels = scales::dollar) +
    facet_wrap(~ continent, ncol = 5) +
    labs(x = "Year", y = "GDP per capita", title = "GDP per capita on Five Continents")

glimpse(gss_sm)

p <- ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) + geom_smooth() + facet_grid(sex ~ race)

p <- ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) + geom_smooth() + facet_grid(sex ~ race + degree)

```
```{r}

p <- ggplot(data = gss_sm, mapping = aes(x = bigregion))
p + geom_bar()

p <- ggplot(data = gss_sm, mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop..))

p <- ggplot(data = gss_sm, mapping = aes(x = bigregion))
p + geom_bar(mapping = aes(y = ..prop.., group = 1))

table(gss_sm$religion)

p <- ggplot(data = gss_sm, mapping = aes(x = religion, color = religion))
p + geom_bar()

p <- ggplot(data = gss_sm, mapping = aes(x = religion, fill = religion))
p + geom_bar() + guides(fill = FALSE)

p <- ggplot(data = gss_sm, mapping = aes(x = bigregion, fill = religion))
p + geom_bar()

p <- ggplot(data = gss_sm, mapping = aes(x = bigregion, fill = religion))
p + geom_bar(position = "fill")

p <- ggplot(data = gss_sm, mapping = aes(x = bigregion, fill = religion))
p + geom_bar(position = "dodge", mapping = aes(y = ..prop..))

p <- ggplot(data = gss_sm, mapping = aes(x = bigregion, fill = religion))
p + geom_bar(position = "dodge", mapping = aes(y = ..prop.., group = religion))

p <- ggplot(data = gss_sm, mapping = aes(x = religion))
p + geom_bar(position = "dodge", mapping = aes(y = ..prop.., group = bigregion)) + facet_wrap(~ bigregion, ncol = 2)

```
```{r}

p <- ggplot(data = midwest, mapping = aes(x = area))
p + geom_histogram()

p <- ggplot(data = midwest, mapping = aes(x = area))
p + geom_histogram(bins = 10)

oh_wi <- c("OH", "WI")
p <- ggplot(data = subset(midwest, subset = state %in% oh_wi), mapping = aes(x = percollege, fill = state))
p + geom_histogram(alpha = 0.4, bins = 20)

p <- ggplot(data = midwest, mapping = aes(x = area))
p + geom_density()

p <- ggplot(data = midwest, mapping = aes(x = area, fill = state, color = state))
p + geom_density(alpha = 0.3)

p <- ggplot(data = midwest, mapping = aes(x = area, color = state))
p + geom_line(stat="density")

p <- ggplot(data = subset(midwest, subset = state %in% oh_wi), mapping = aes(x = area, fill = state, color = state))
p + geom_density(alpha = 0.3, mapping = aes(y = ..scaled..))

p <- ggplot(data = subset(midwest, subset = state %in% oh_wi), mapping = aes(x = area, fill = state, color = state))
p + geom_density(alpha = 0.3, mapping = aes(y = ..count..))

```
```{r}

titanic

p <- ggplot(data = titanic, mapping = aes(x = fate, y = percent, fill = sex))
p + geom_bar(position = "dodge", stat = "identity") + theme(legend.position = "top")

oecd_sum

p <- ggplot(data = oecd_sum, mapping = aes(x = year, y = diff, fill = hi_lo))

p + geom_col() + guides(fill = FALSE) +
    labs(x = NULL, y = "Difference in Years", title = "The US Life Expectancy Gap",
         subtitle = "Difference between US and OECD average life expectancies, 1960-2015",
         caption = "Data: OECD. After a chart by Christopher Ingraham, Washington Post, December 27th 2017.")

p <- ggplot(data = oecd_sum, mapping = aes(x = year, y = diff, fill = hi_lo))

p + geom_bar(position = "identity", stat = "identity") + guides(fill = FALSE) +
    labs(x = NULL, y = "Difference in Years", title = "The US Life Expectancy Gap",
         subtitle = "Difference between US and OECD average life expectancies, 1960-2015",
         caption = "Data: OECD. After a chart by Christopher Ingraham, Washington Post, December 27th 2017.")

```
```{r, fig.width = 10, fig.height = 8}

p <- ggplot(data = gapminder, mapping = aes(x = pop, y = gdpPercap))
p + geom_point(mapping = aes(color = continent)) + scale_x_log10() + scale_y_log10(labels = scales::dollar) + facet_wrap(~ year)

```
```{r}

p <- ggplot(data = gapminder, mapping = aes(x = year, y = pop))
p_out <- p + geom_line(aes(group = country)) + scale_y_log10() + facet_wrap(~country)

ggsave(here("figures", "pop_vs_year_countries.pdf"), plot = p_out, 
       height = 12, width = 20, units = "in")

p <- ggplot(data = gapminder, mapping = aes(x = year, y = gdpPercap))
p_out <- p + geom_line(aes(group = country)) + scale_y_log10() + facet_wrap(~country)

ggsave(here("figures", "dgpPercap_vs_year_countries.pdf"), plot = p_out, 
       height = 12, width = 20, units = "in")

```

```{r}

p <- ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) + geom_smooth() + facet_grid(sex ~ race)

p <- ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) + geom_smooth() + facet_grid(~ sex + race)

p <- ggplot(data = gss_sm, mapping = aes(x = age, y = childs))
p + geom_point(alpha = 0.2) + geom_smooth() + facet_wrap(~ sex + race)

```
```{r}

p <- ggplot(data = midwest, mapping = aes(x = area))
p + geom_freqpoly(size = 1.5,)

p <- ggplot(data = midwest, mapping = aes(x = area))
p + geom_freqpoly(size = 1.5, bins = 10)

oh_wi <- c("OH", "WI")
p <- ggplot(data = subset(midwest, subset = state %in% oh_wi), mapping = aes(x = percollege, color = state))
p + geom_freqpoly(alpha = 0.4, size = 1.5, bins = 20)

p <- ggplot(data = gapminder, mapping = aes(x = lifeExp, y = gdpPercap))
p + geom_bin2d(bins = c(20, 50))  # binwidth = 5

p <- ggplot(data = midwest, mapping = aes(x = percollege, y = percbelowpoverty))
p + geom_density_2d()

p <- ggplot(data = midwest, mapping = aes(x = percollege, y = percbelowpoverty))
p + geom_density_2d() + geom_point()

```

## Graph Tables, Make Labels, Add Notes

```{r}

rel_by_region <- gss_sm %>% 
    group_by(bigregion, religion) %>%
    summarize(N = n()) %>%
    mutate(freq = N / sum(N),
           pct = round(freq*100, 0))

rel_by_region

rel_by_region %>% group_by(bigregion) %>% summarize(total = sum(pct))

p <- ggplot(rel_by_region, aes(x = bigregion, y = pct, fill = religion))

p + geom_col(position = "dodge") +
    labs(x = "Region", y = "Percent", fill = "Religion") +
    theme(legend.position = "top")

p <- ggplot(rel_by_region, aes(x = bigregion, y = pct, fill = religion))

p + geom_col(position = "dodge2") +
    labs(x = "Region", y = "Percent", fill = "Religion") +
    theme(legend.position = "top")

p <- ggplot(rel_by_region, aes(x = religion, y = pct, fill = religion))

p + geom_col(position = "dodge2") +
    labs(x = NULL, y = "Percent", fill = "Religion") +
    guides(fill = FALSE) +
    coord_flip() +
    facet_grid(~ bigregion)

```
```{r}

head(organdata)
organdata %>% select(1:6) %>% sample_n(size = 10)

p <- ggplot(data = organdata, mapping = aes(x = year, y = donors))
p + geom_point()

p <- ggplot(data = organdata, mapping = aes(x = year, y = donors))
p + geom_line(aes(group = country)) + facet_wrap(~ country)

p <- ggplot(data = organdata, mapping = aes(x = country, y = donors))
p + geom_boxplot()      # outlier.shape, outlier.color

p <- ggplot(data = organdata, mapping = aes(x = country, y = donors))
p + geom_boxplot() + coord_flip()

p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm = TRUE), y = donors))
p + geom_boxplot() + labs(x = NULL) + coord_flip()

p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm = TRUE), y = donors))
p + geom_violin() + labs(x = NULL) + coord_flip()

p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, sd, na.rm = TRUE), y = donors))
p + geom_boxplot() + labs(x = NULL) + coord_flip()

p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, median, na.rm = TRUE), y = donors))
p + geom_boxplot() + labs(x = NULL) + coord_flip()

p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm = TRUE), y = donors, fill = world))
p + geom_boxplot() + labs(x = NULL) + coord_flip() + 
    theme(legend.position = "top")

p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm = TRUE), y = donors, color = world))
p + geom_point() + labs(x = NULL) + coord_flip() + 
    theme(legend.position = "top")

p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm = TRUE), y = donors, color = world))
p + geom_jitter() + labs(x = NULL) + coord_flip() + 
    theme(legend.position = "top")

p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm = TRUE), y = donors, color = world))
p + geom_jitter(position = position_jitter(width = 0.15)) + labs(x = NULL) + coord_flip() + 
    theme(legend.position = "top")

p <- ggplot(data = organdata, mapping = aes(x = reorder(country, donors, na.rm = TRUE), y = donors, color = world))
p + geom_jitter(position = position_jitter(height = 0.15)) + labs(x = NULL) + coord_flip() + 
    theme(legend.position = "top")

```
```{r}

glimpse(organdata)

by_country <- organdata %>% group_by(consent_law, country) %>%
    summarize(donors_mean = mean(donors, na.rm = TRUE),
              donors_sd = sd(donors, na.rm = TRUE),
              gdp_mean = mean(gdp, na.rm = TRUE),
              health_mean = mean(health, na.rm = TRUE),
              roads_mean = mean(roads, na.rm = TRUE),
              cerebvas_mean = mean(cerebvas, na.rm = TRUE))

by_country

by_year <- organdata %>% group_by(consent_law, year) %>%
    summarize(donors_mean = mean(donors, na.rm = TRUE),
              donors_sd = sd(donors, na.rm = TRUE),
              gdp_mean = mean(gdp, na.rm = TRUE),
              health_mean = mean(health, na.rm = TRUE),
              roads_mean = mean(roads, na.rm = TRUE),
              cerebvas_mean = mean(cerebvas, na.rm = TRUE))

by_year

by_country <- organdata %>% group_by(consent_law, country) %>%
    summarize_if(is.numeric, funs(mean, sd), na.rm = TRUE) %>%
    ungroup()

by_country

p <- ggplot(data = by_country, mapping = aes(x = donors_mean, y = reorder(country, donors_mean), color = consent_law))

p + geom_point(size = 3) +
    labs(x = "Donor Procurement Rate", y = NULL, color = "Consent Law") +
    theme(legend.position = "top")

p <- ggplot(data = by_country, mapping = aes(x = donors_mean, y = reorder(country, donors_mean)))

p + geom_point(size = 3) +
    facet_wrap(~ consent_law, scales = "free_y", ncol = 1) +
    labs(x = "Donor Procurement Rate", y = NULL) 

p <- ggplot(data = by_country, mapping = aes(x = reorder(country, donors_mean), y = donors_mean))

p + geom_pointrange(mapping = aes(ymin = donors_mean - donors_sd, ymax = donors_mean + donors_sd)) +
    labs(x = NULL, y = "Donor Procurement Rate") + coord_flip()

```
```{r, fig.width = 10, fig.height = 8}

p <- ggplot(data = by_country, mapping = aes(x = roads_mean, y = donors_mean))
p + geom_point() + geom_text(mapping = aes(label = country))

p <- ggplot(data = by_country, mapping = aes(x = roads_mean, y = donors_mean))
p + geom_point() + geom_text(mapping = aes(label = country), hjust = 0)

p <- ggplot(data = by_country, mapping = aes(x = roads_mean, y = donors_mean))
p + geom_point() + geom_text(mapping = aes(label = country), hjust = 1)

library(ggrepel)

elections_historic %>% select(2:7)

p_title <- "Presidential Elections: Popular & Electoral College Margins"
p_subtitle <- "1824-2016"
p_caption <- "Data for 2016 are provisional."
x_label <- "Winner's share of Popular Vote"
y_label <- "Winner's share of Electoral College Votes"

p <- ggplot(elections_historic, aes(x = popular_pct, y = ec_pct, label = winner_label))

p + geom_hline(yintercept = 0.5, size = 1.4, color = "gray80") +
    geom_vline(xintercept = 0.5, size = 1.4, color = "gray80") +
    geom_abline(color = "gray80") +
    geom_point() +
    geom_text_repel() +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(labels = scales::percent) +
    labs(x = x_label, y = y_label, title = p_title, subtitle = p_subtitle, caption = p_caption)

```
```{r}

library(ggrepel)

p <- ggplot(data = by_country, mapping = aes(x = gdp_mean, y = health_mean))

p + geom_point() +
    geom_text_repel(data = subset(by_country, gdp_mean > 25000),
                    mapping = aes(label = country))

p <- ggplot(data = by_country, mapping = aes(x = gdp_mean, y = health_mean))

p + geom_point() +
    geom_text_repel(data = subset(by_country, gdp_mean > 25000 | health_mean < 1500 | country %in% "Belgium"),
                    mapping = aes(label = country))

organdata$ind <- organdata$ccode %in% c("Ita", "Spa") & organdata$year > 1998

p <- ggplot(data = organdata, mapping = aes(x = roads, y = donors, color = ind))

p + geom_point() +
    geom_text_repel(data = subset(organdata, ind), mapping = aes(label = ccode)) +
    guides(label = FALSE, color = FALSE)

p <- ggplot(data = organdata, mapping = aes(x = roads, y = donors))

p + geom_point(aes(color = ind)) +
    geom_text_repel(data = subset(organdata, ind), mapping = aes(label = ccode)) +
    guides(label = FALSE, color = FALSE)

p <- ggplot(data = organdata, mapping = aes(x = roads, y = donors))

p + geom_point() +
    annotate(geom = "text", x = 91, y = 33, label = "A suprisingly high \n recovery rate.", hjust = 0)

p <- ggplot(data = organdata, mapping = aes(x = roads, y = donors))

p + geom_point() +
    annotate(geom = "rect", xmin = 125, xmax = 155, ymin = 30, ymax = 35, fill = "red", alpha = 0.2) +
    annotate(geom = "text", x = 157, y = 33, label = "A suprisingly high \n recovery rate.", hjust = 0)

```
```{r}

# scale_<MAPPING>_<KIND>()

p <- ggplot(data = organdata, mapping = aes(x = roads, y = donors, color = world))
p + geom_point()

p <- ggplot(data = organdata, mapping = aes(x = roads, y = donors, color = world))

p + geom_point() +
    scale_x_log10() +
    scale_y_continuous(breaks = c(5, 15, 25), labels = c("Five", "Fifteen", "Twenty Five"))

p <- ggplot(data = organdata, mapping = aes(x = roads, y = donors, color = world))

p + geom_point() +
    scale_color_discrete(labels = c("Corporatist", "Liberal", "Social Democratic", "Unclassified")) +
    labs(x = "Road Deaths", y = "Donor Procurement", color = "Welfare State")

p <- ggplot(data = organdata, mapping = aes(x = roads, y = donors, color = world))

p + geom_point() +
    labs(x = "Road Deaths", y = "Donor Procurement") +
    guides(color = FALSE)

```
```{r}

library(ggrepel)
glimpse(elections_historic)

p_title <- "Presidential Elections: Popular & Electoral College Margins"
p_subtitle <- "1824-2016"
p_caption <- "Data for 2016 are provisional."
x_label <- "Winner's share of Popular Vote"
y_label <- "Winner's share of Electoral College Votes"

p <- ggplot(elections_historic, aes(x = popular_pct, y = ec_pct, label = winner_label))

p + geom_hline(yintercept = 0.5, size = 1.4, color = "gray80") +
    geom_vline(xintercept = 0.5, size = 1.4, color = "gray80") +
    geom_point(aes(color = ru_part)) +
    geom_text_repel(data = subset(elections_historic, year >= 1992), 
                    mapping = aes(color = ru_part),
                    show.legend = FALSE) +
    annotate(geom = "rect", xmin = 0.3, xmax = 0.5, ymin = 0.5, ymax = 1, fill = "steelblue", alpha = 0.1) +
    scale_x_continuous(labels = scales::percent) +
    scale_y_continuous(labels = scales::percent) +
    labs(x = x_label, y = y_label, title = p_title, subtitle = p_subtitle, caption = p_caption)

p <- ggplot(data = elections_historic, mapping = aes(x = reorder(year, popular_pct), y = popular_pct))

p + geom_point(size = 3, aes(color = ru_part)) +
    labs(x = "Donor Procurement Rate", y = NULL) +
    scale_y_continuous(labels = scales::percent) +
    coord_flip()

```
```{r}
 
life_by_continent <- gapminder %>% filter(year == 2007) %>% group_by(continent) %>% summarize(m = mean(lifeExp))
life_by_continent

p <- ggplot(data = life_by_continent, mapping = aes(x = reorder(continent, m), y = m))

p + geom_col() + 
    labs(x = NULL, y = "Life expectancy in years, 2007") +
    coord_flip()

p <- ggplot(data = life_by_continent, mapping = aes(x = reorder(continent, m), y = m))

p + geom_point(size = 5) + 
    labs(x = NULL, y = "Life expectancy in years, 2007") +
    coord_flip()

degree_by_race <- gss_sm %>% group_by(race, degree) %>% summarize(N = n()) %>% mutate(pct = round(N/sum(N)*100, 0))
degree_by_race
degree_by_race %>% group_by(race) %>% summarize(total = sum(pct))

degree_by_sex <- gss_sm %>% group_by(sex, degree) %>% summarize(N = n()) %>% mutate(pct = round(N/sum(N)*100, 0))
degree_by_sex
degree_by_sex %>% group_by(sex) %>% summarize(total = sum(pct))

degree_by_region <- gss_sm %>% group_by(region, degree) %>% summarize(N = n()) %>% mutate(pct = round(N/sum(N)*100, 0))
degree_by_region
degree_by_region %>% group_by(region) %>% summarize(total = sum(pct))

gss_sm %>% group_by(degree) %>% summarize(mean = mean(childs, na.rm = TRUE), median = median(childs, na.rm = TRUE))

```
```{r}

str(gapminder)

p <- ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp, group = year))
p + geom_boxplot()

p <- ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp, group = year))
p + geom_boxplot() + facet_wrap(~ continent)

p <- ggplot(data = gapminder, mapping = aes(x = year, y = pop, group = year))
p + geom_boxplot() + scale_y_log10() + facet_wrap(~ continent)

life_by_year <- gapminder %>% group_by(continent, year)
life_by_year    # this produce improper result

p <- ggplot(data = life_by_year, mapping = aes(x = year, y = lifeExp), group = year)
p + geom_boxplot() + facet_wrap(~ continent)

p <- ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp, group = year))
p + geom_boxplot(notch = TRUE) + facet_wrap(~ continent)

p <- ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp, group = year))
p + geom_boxplot(notch = TRUE, varwidth = TRUE)

p <- ggplot(data = gapminder, mapping = aes(x = year, y = lifeExp, group = year))
p + geom_violin()

```
```{r}

by_country <- organdata %>% group_by(consent_law, country) %>%
    summarize_if(is.numeric, funs(mean, sd), na.rm = TRUE) %>%
    ungroup()

p <- ggplot(data = by_country, mapping = aes(x = reorder(country, donors_mean), y = donors_mean))

p + geom_pointrange(mapping = aes(ymin = donors_mean - donors_sd, ymax = donors_mean + donors_sd)) +
    labs(x = NULL, y = "Donor Procurement Rate") + coord_flip()

p + geom_linerange(mapping = aes(ymin = donors_mean - donors_sd, ymax = donors_mean + donors_sd)) +
    labs(x = NULL, y = "Donor Procurement Rate") + coord_flip()

p + geom_crossbar(mapping = aes(ymin = donors_mean - donors_sd, ymax = donors_mean + donors_sd)) +
    labs(x = NULL, y = "Donor Procurement Rate") + coord_flip()

p + geom_errorbar(mapping = aes(ymin = donors_mean - donors_sd, ymax = donors_mean + donors_sd)) +
    labs(x = NULL, y = "Donor Procurement Rate") + coord_flip()

```

## Work with Models

```{r}

p <- ggplot(data = gapminder, mapping = aes(x = log(gdpPercap), y = lifeExp))

p + geom_point(alpha = 0.1) +
    geom_smooth(color = "tomato", fill = "tomato", method = MASS::rlm) +
    geom_smooth(color = "steelblue", fill = "steelblue", method = "lm")
    
p + geom_point(alpha = 0.1) +
    geom_smooth(color = "tomato", size = 1.2, method = "lm", formula = y ~ splines::bs(x, 3), se = FALSE)

p + geom_point(alpha = 0.1) +
    geom_quantile(color = "tomato", size = 1.2, method = "rqss", lambda = 1, quantiles = c(0.2, 0.5, 0.85))

model_colors <- RColorBrewer::brewer.pal(3, "Set1")
model_colors

p0 <- ggplot(data = gapminder, mapping = aes(x = log(gdpPercap), y = lifeExp))

p1 <- p0 + geom_point(alpha = 0.2) +
    geom_smooth(method = "lm", aes(color = "OLS", fill = "OLS")) +
    geom_smooth(method = "lm", formula = y ~ splines::bs(x, df = 3), aes(color = "Cubic Spline", fill = "Cubic Spline")) +
    geom_smooth(method = "loess", aes(color = "LOESS", fill = "LOESS")) 

p1 + scale_color_manual(name = "Models", values = model_colors) +
     scale_fill_manual(name = "Models", values = model_colors) +
     theme(legend.position = "top")

```
```{r}

gapminder
str(gapminder)
summary(gapminder)

out <- lm(formula = lifeExp ~ gdpPercap + pop + continent, data = gapminder)
summary(out)
str(out)

out$coefficients
#out$residuals
#out$fitted.values
out$df.residual
#out$model

min_gdp <- min(gapminder$gdpPercap)
max_gdp <- max(gapminder$gdpPercap)
med_pop <- median(gapminder$pop)

pred_df <- expand.grid(gdpPercap = seq(from = min_gdp, to = max_gdp, length.out = 100),
                       pop = med_pop, continent = c("Africa", "Americas", "Asia", "Europe", "Oceania"))

dim(pred_df)
head(pred_df)

pred_out <- predict(object = out, newdata = pred_df, interval = "predict")
head(pred_out)

pred_df <- cbind(pred_df, pred_out)
head(pred_df)

p <- ggplot(data = subset(pred_df, continent %in% c("Europe", "Africa")),
            aes(x = gdpPercap, y = fit, ymin = lwr, ymax = upr,
                color = continent, fill = continent, group = continent))

p + geom_point(data = subset(gapminder, continent %in% c("Europe", "Africa")),
               aes(x = gdpPercap, y = lifeExp, color = continent),
               alpha = 0.5, inherit.aes = FALSE) +
    geom_line() +
    geom_ribbon(alpha = 0.2, color = FALSE) +
    scale_x_log10(labels = scales::dollar)

```
```{r}

library(broom)

out_comp <- tidy(out)
out_comp %>% round_df()

p <- ggplot(out_comp, mapping = aes(x = term, y = estimate))
p + geom_point() + coord_flip()

out_conf <- tidy(out, conf.int = TRUE)
out_conf %>% round_df()

out_conf <- subset(out_conf, term %nin% "(Intercept)")
out_conf$nicelabs <- prefix_strip(out_conf$term, "continent")

p <- ggplot(out_conf, mapping = aes(x = reorder(nicelabs, estimate),
                                    y = estimate, ymin = conf.low, ymax = conf.high))
p + geom_pointrange() + coord_flip() + labs(x = NULL, y = "OLS Estimate")

out_aug <- augment(out)
head(out_aug) %>% round_df()

out_aug <- augment(out, data = gapminder)
head(out_aug) %>% round_df()

p <- ggplot(data = out_aug, mapping = aes(x = .fitted, y = .resid))
p + geom_point()

glance(out) %>% round_df()

library(survival)

out_cph <- coxph(Surv(time, status) ~ age + sex, data = lung)
out_surv <- survfit(out_cph)

summary(out_cph)
#summary(out_surv)

out_tidy <- tidy(out_surv)

p <- ggplot(data = out_tidy, mapping = aes(time, estimate))
p + geom_line() + geom_ribbon(mapping = aes(ymin = conf.low, ymax = conf.high), alpha = 0.2)

```
```{r}

eu77 <- gapminder %>% filter(continent == "Europe", year == 1977)
fit <- lm(lifeExp ~ log(gdpPercap), data = eu77)
summary(fit)

out_le <- gapminder %>% group_by(continent, year) %>% nest()
out_le

out_le %>% filter(continent == "Europe" & year == 1977) %>% unnest()

fit_ols <- function(df) {
    lm(lifeExp ~ log(gdpPercap), data = df)
}

fit_ols
fit_ols(df = gapminder)
summary(fit_ols(gapminder))

out_le <- gapminder %>% group_by(continent, year) %>% nest() %>%
    mutate(model = map(data, fit_ols))
        
out_le

out_tidy <- gapminder %>% group_by(continent, year) %>% nest() %>%
    mutate(model = map(data, fit_ols), tidied = map(model, tidy)) %>%
    select(!(data:model)) %>%
    unnest(tidied) %>%
    filter(term %nin% "(Intercept)" & continent %nin% "Oceania") %>%
    ungroup()

out_tidy %>% sample_n(5)

p <- ggplot(data = out_tidy, mapping = aes(x = year, y = estimate,
                                           ymin = estimate - 2*std.error, ymax = estimate + 2*std.error,
                                           group = continent, color = continent))

p + geom_pointrange(position = position_dodge(width = 1)) +
    scale_x_continuous(breaks = unique(gapminder$year)) +
    theme(legend.position = "top") +
    labs(x = "Year", y = "Estimate", color = "Continent")

```
```{r}

library(margins)

gss_sm$polviews_m <- relevel(gss_sm$polviews, ref = "Moderate")
out_bo <- glm(obama ~ polviews_m + sex*race, family = "binomial", data = gss_sm)
summary(out_bo)

bo_m <- margins(out_bo)
summary(bo_m)

plot(bo_m)

bo_gg <- as_tibble(summary(bo_m))
prefixes <- c("polviews_m", "sex")
bo_gg$factor <- prefix_strip(bo_gg$factor, prefixes)
bo_gg$factor <- prefix_replace(bo_gg$factor, "race", "Race: ")

bo_gg %>% select(factor, AME, lower, upper)

p <- ggplot(data = bo_gg, aes(x = reorder(factor, AME),
                              y = AME, ymin = lower, ymax = upper))
p + geom_hline(yintercept = 0, color = "gray80") +
    geom_pointrange() + coord_flip() +
    labs(x = NULL, y = "Average Marginal Effect")

cplot(out_bo, "sex")

pv_cp <- cplot(out_bo, x = "sex", draw = FALSE)

p <- ggplot(data = pv_cp, aes(x = reorder(xvals, yvals),
                              y = yvals, ymin = lower, ymax = upper))
p + geom_hline(yintercept = 0, color = "gray80") +
    geom_pointrange() + coord_flip() +
    labs(x = NULL, y = "Conditional Effect")

```
```{r, fig.width = 6, fig.height = 10}

library(survey)
library(srvyr)

options(survey.lonely.psu = "adjust")
options(na.action = "na.pass")

gss_wt <- subset(gss_lon, year > 1974) %>%
    mutate(stratvar = interaction(year, vstrat)) %>%
    as_survey_design(ids = vpsu, strata = stratvar, weights = wtssall, nest = TRUE)

out_grp <- gss_wt %>%
    filter(year %in% seq(1976, 2016, by = 4), !is.na(degree)) %>%
    group_by(year, race, degree) %>%
    summarize(prop = survey_mean(na.rm = TRUE))

out_grp

out_mrg <- gss_wt %>%
    filter(year %in% seq(1976, 2016, by = 4)) %>%
    mutate(racedeg = interaction(race, degree)) %>%
    group_by(year, racedeg) %>%
    summarize(prop = survey_mean(na.rm = TRUE))

out_mrg

out_mrg <- gss_wt %>%
    filter(year %in% seq(1976, 2016, by = 4)) %>%
    mutate(racedeg = interaction(race, degree)) %>%
    group_by(year, racedeg) %>%
    summarize(prop = survey_mean(na.rm = TRUE)) %>%
    separate(racedeg, sep = "\\.", into = c("race", "degree"))

out_mrg

p <- ggplot(data = subset(out_grp, race %nin% "Other"),
            mapping = aes(x = degree, y = prop, ymin = prop - 2*prop_se, ymax = prop + 2*prop_se,
                          fill = race, color = race))

dodge <- position_dodge(width = 0.9)

p + geom_col(position = dodge, alpha = 0.2) +
    geom_errorbar(position = dodge, width = 0.2) +
    scale_x_discrete(labels = scales::wrap_format(10)) +
    scale_y_continuous(labels = scales::percent) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer(type = "qual", palette = "Dark2") +
    labs(title = "Educational Attainment by Race",
         subtitle = "GSS 1976-2016",
         fill = "Race", color = "Race", x = NULL, y = "Percent") +
    facet_wrap(~ year, ncol = 2) +
    theme(legend.position = "top")

p <- ggplot(data = subset(out_grp, race %nin% "Other"),
            mapping = aes(x = year, y = prop, ymin = prop - 2*prop_se, ymax = prop + 2*prop_se,
                          fill = race, color = race))

p + geom_ribbon(alpha = 0.3, aes(color = NULL)) +
    geom_line() +
    scale_y_continuous(labels = scales::percent) +
    scale_color_brewer(type = "qual", palette = "Dark2") +
    scale_fill_brewer(type = "qual", palette = "Dark2") +
    labs(title = "Educational Attainment by Race",
         subtitle = "GSS 1976-2016",
         fill = "Race", color = "Race", x = NULL, y = "Percent") +
    facet_wrap(~ degree, ncol = 1) +
    theme(legend.position = "top")

```
```{r}

out <- lm(formula = lifeExp ~ log(gdpPercap) + pop + continent, data = gapminder)
plot(out, which = c(1, 2), ask = FALSE)

#install.packages("ggfortify")   # for using autoplot()

library(coefplot)

out <- lm(formula = lifeExp ~ log(gdpPercap) + log(pop) + continent, data = gapminder)
coefplot(out, sort = "magnitude", intercept = FALSE)

#install.packages("infer")

library(GGally)

organdata_sm <- organdata %>% select(donors, pop_dens, pubhealth, roads, consent_law)

ggpairs(data = organdata_sm, mapping = aes(color = consent_law),
        upper = list(continuous = wrap("density"), combo = wrap("box_no_facet")),
        lower = list(continuous = wrap("points"), combo = wrap("dot_no_facet")))

```

## Draw Maps

```{r, fig.width = 4, fig.height = 8}

election %>% select(state, total_vote, r_points, pct_trump, party, census) %>%
    sample_n(5)

# Hex color codes for Dem Blue and Rep Red
party_colors <- c("#2E74C0", "#CB454A")

p0 <- ggplot(data = subset(election, st %nin% "DC"),
             mapping = aes(x = r_points, y = reorder(state, r_points), color = party))

p1 <- p0 + geom_vline(xintercept = 0, color = "gray30") +
    geom_point(size = 2)

p2 <- p1 + scale_color_manual(values = party_colors)

p3 <- p2 + scale_x_continuous(breaks = c(-30, -20, -10, 0, 10, 20, 30, 40),
                              labels = c("30\n(Clinton)", "20", "10", "0", "10", "20", "30", "40\n(Trump)"))

p3 + facet_wrap(~ census, ncol = 1, scales = "free_y") +
    guides(color = FALSE) + labs(x = "Point Margin", y = NULL) +
    theme(axis.text = element_text(size = 8))

```
```{r}

library(maps)
us_states <- map_data("state")
head(us_states)
dim(us_states)

p <- ggplot(data = us_states, mapping = aes(x = long, y = lat, group = group))
p + geom_polygon(fill = "white", color = "black")

p <- ggplot(data = us_states, mapping = aes(x = long, y = lat, group = group, fill = region))
p + geom_polygon(color = "gray90", size = 0.1) + guides(fill = FALSE)

p <- ggplot(data = us_states, mapping = aes(x = long, y = lat, group = group, fill = region))
p + geom_polygon(color = "gray90", size = 0.1) + 
    coord_map(projection = "albers", lat0 = 39, lat1 = 45) + guides(fill = FALSE)

theme_map <- function(base_size = 9, base_family = NULL) {
    
    require(grid)
    theme_bw(base_size = base_size, base_family = base_family) %+replace%
        theme(axis.line = element_blank(),
              axis.text = element_blank(),
              axis.ticks = element_blank(),
              axis.title = element_blank(),
              panel.background = element_blank(),
              panel.border = element_blank(),
              panel.grid = element_blank(),
              panel.spacing = unit(0, "lines"),
              plot.background = element_blank(),
              legend.justification = c(0,0),
              legend.position = c(0,0))
    
}

```
```{r}

election$region <- tolower(election$state)
us_states_elec <- left_join(us_states, election, by = "region") # instead of: merge(us_states, election, sort = FALSE)
head(us_states_elec)

p0 <- ggplot(data = us_states_elec, mapping = aes(x = long, y = lat, group = group, fill = party))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) + 
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)

p2 <- p1 + scale_fill_manual(values = party_colors) +
    labs(title = "Election Results 2016", fill = NULL)
p2 + theme_map()

p0 <- ggplot(data = us_states_elec, mapping = aes(x = long, y = lat, group = group, fill = pct_trump))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) + 
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)
p1 + labs(title = "Trump vote") + theme_map() + labs(fill = "Percent")

p2 <- p1 + scale_fill_gradient(low = "white", high = "#CB454A") +
    labs(title = "Election Results 2016", fill = NULL)
p2 + labs(title = "Trump vote") + theme_map() + labs(fill = "Percent")

p0 <- ggplot(data = us_states_elec, mapping = aes(x = long, y = lat, group = group, fill = d_points))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) + 
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)

p2 <- p1 + scale_fill_gradient2() + labs(title = "Winning margins") 
p2 + theme_map() + labs(fill = "Percent")

p3 <- p1 + scale_fill_gradient2(low = "red", mid = scales::muted("purple"), high = "blue",
                                breaks = c(-25, 0, 25, 50, 75)) + 
    labs(title = "Winning margins") 
p3 + theme_map() + labs(fill = "Percent")

p0 <- ggplot(data = subset(us_states_elec, region %nin% "district of columbia"),
             mapping = aes(x = long, y = lat, group = group, fill = d_points))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.1) + 
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)

p2 <- p1 + scale_fill_gradient2(low = "red", mid = scales::muted("purple"), high = "blue") + 
    labs(title = "Winning margins") 
p2 + theme_map() + labs(fill = "Percent")

```
```{r}

county_map %>% sample_n(5)

county_data %>% select(id, name, state, pop_dens, pct_black) %>% sample_n(5)
head(county_data)

county_full <- left_join(county_map, county_data, by = "id")

p <- ggplot(data = county_full, mapping = aes(x = long, y = lat, fill = pop_dens, group = group))

p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

p2 <- p1 + scale_fill_brewer(palette = "Blues",
                             labels = c("0-10", "10-50", "50-100", "100-500", "500-1,000", "1,000-5,000", ">5,000"))
p2 + labs(fill = "Population per\nsquare mile") + theme_map() +
    guides(fill = guide_legend(nrow = 1)) + theme(legend.position = "bottom")

p <- ggplot(data = county_full, mapping = aes(x = long, y = lat, fill = pct_black, group = group))

p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

p2 <- p1 + scale_fill_brewer(palette = "Greens")
p2 + labs(fill = "US Population, Percent Black") + theme_map() +
    guides(fill = guide_legend(nrow = 1)) + theme(legend.position = "bottom")

orange_pal <- RColorBrewer::brewer.pal(n = 6, name = "Oranges")
orange_pal

orange_rev <- rev(orange_pal)
orange_rev

gun_p <- ggplot(data = county_full, mapping = aes(x = long, y = lat, fill = su_gun6, group = group))

gun_p1 <- gun_p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

gun_p2 <- gun_p1 + scale_fill_manual(values = orange_pal)
gun_p2 + labs(title = "Gun-Related Suicides, 1999-2015", fill = "Rate per 100,000 pop") + 
    theme_map() + theme(legend.position = "bottom")

pop_p <- ggplot(data = county_full, mapping = aes(x = long, y = lat, fill = pop_dens6, group = group))

pop_p1 <- pop_p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

pop_p2 <- pop_p1 + scale_fill_manual(values = orange_rev)
pop_p2 + labs(title = "Reverse-coded Population Density", fill = "People per square mile") + 
    theme_map() + theme(legend.position = "bottom")

```
```{r}

library(statebins)

p0 <- ggplot(data = election, mapping = aes(state = state, fill = pct_trump))

p0 + geom_statebins() + coord_equal() +
     scale_fill_distiller(palette = "Reds", direction = 1) +
     labs(fill = "Percent Trump") +
     theme_statebins(legend_position = "top")

p0 <- ggplot(data = subset(election, st %nin% "DC"), mapping = aes(state = state, fill = pct_clinton))

p0 + geom_statebins(lbl_size = 4, light_lbl = "black") + coord_equal() +
     scale_fill_distiller(palette = "Blues", direction = 1) +
     labs(fill = "Percent Clinton") +
     theme_statebins(legend_position = "top")

# Hex color codes for Dem Blue and Rep Red
party_colors <- c("#CB454A", "#2E74C0")

election <- election %>% mutate(color = recode(party, Republican = party_colors[2], Democratic = party_colors[1]))

p0 <- ggplot(data = election, mapping = aes(state = state, fill = color))

p0 + geom_statebins(lbl_size = 4) + coord_equal() +
     scale_fill_manual(values = party_colors, labels = c("Trump", "Clinton")) +
     labs(fill = "Winner") + 
     theme_statebins(legend_position = "right")

election$category <- cut(election$pct_trump, breaks=c(4, 21, 37, 53, 70), 
                         labels = c("4-21", "21-37", "37-53", "53-70"))

model_colors <- RColorBrewer::brewer.pal(n = 4, name = "Reds")

p0 <- ggplot(data = election, mapping = aes(state = state, fill = category))

p0 + geom_statebins(lbl_size = 4) + coord_equal() +
     scale_fill_manual(values = model_colors, labels = c("4-21", "21-37", "37-53", "53-70")) +
     labs(fill = "Percent Trump") +
     theme_statebins(legend_position = "top")

```
```{r, fig.width = 10, fig.height = 12}

opiates
opiates$region <- tolower(opiates$state)
opiates_map <- left_join(us_states, opiates, by = "region")

library(viridis)

p0 <- ggplot(data = subset(opiates_map, year > 1999),
             mapping = aes(x = long, y = lat, group = group, fill = adjusted))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.05) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)

p2 <- p1 + scale_fill_viridis_c(option = "plasma")

p2 + theme_map() + facet_wrap(~ year, ncol = 3) +
    theme(legend.position = "bottom", strip.background = element_blank()) +
    labs(fill = "Death rate per 100,000 population",
         title = "Opiate Related Deaths by State, 2000-2014")

opiates_map$category <- cut_interval(opiates_map$adjusted, n = 5)

p0 <- ggplot(data = subset(opiates_map, year > 1999),
             mapping = aes(x = long, y = lat, group = group, fill = category))

p1 <- p0 + geom_polygon(color = "gray90", size = 0.05) +
    coord_map(projection = "albers", lat0 = 39, lat1 = 45)

p2 <- p1 + scale_fill_viridis_d(option = "plasma")

p2 + theme_map() + facet_wrap(~ year, ncol = 3) +
    theme(legend.position = "bottom", strip.background = element_blank()) +
    labs(fill = "Death rate per 100,000 population",
         title = "Opiate Related Deaths by State, 2000-2014")

```
```{r}

library(ggrepel)

p <- ggplot(data = opiates, mapping = aes(x = year, y = adjusted, group = state))
p + geom_line(color = "gray70")

p0 <- ggplot(data = drop_na(opiates, division_name), mapping = aes(x = year, y = adjusted))

p1 <- p0 + geom_line(color = "gray70", mapping = aes(group = state))

p2 <- p1 + geom_smooth(mapping = aes(group = division_name), se = FALSE)

p3 <- p2 + geom_text_repel(data = subset(opiates, year == max(year) & abbr != "DC"),
                           mapping = aes(x = year, y = adjusted, label = abbr),
                           size = 1.8, segment.color = NA, nudge_x = 30) +
    coord_cartesian(c(min(opiates$year), max(opiates$year)))

p3 + labs(x = NULL, y = "Rate per 100,000 population",
          title = "State-Level Opiate Death Rates by Census Division, 1999-2014") +
    facet_wrap(~ reorder(division_name, - adjusted, na.rm = TRUE), nrow = 3)

```
```{r}

p <- ggplot(data = county_full, mapping = aes(x = long, y = lat, fill = black, group = group))
p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

p2 <- p1 + scale_fill_viridis_c(option = "plasma")
p2 + labs(fill = "US Population, Percent Black") + theme_map() +
    guides(fill = guide_legend(nrow = 1)) + theme(legend.position = "bottom")

p <- ggplot(data = county_full, mapping = aes(x = long, y = lat, fill = pop/land_area, group = group))
p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

p2 <- p1 + scale_fill_viridis_c(option = "inferno")
p2 + labs(fill = "Population per square mile") + theme_map() +
    guides(fill = guide_legend(nrow = 1)) + theme(legend.position = "bottom")

p <- ggplot(data = county_full, mapping = aes(x = long, y = lat, fill = pop/land_area, group = group))
p1 <- p + geom_polygon(color = "gray90", size = 0.05) + coord_equal()

p2 <- p1 + scale_fill_viridis_c(option = "inferno", breaks = c(1, 10, 100, 1000, 1000, 10000, 70000),
                                trans = scales::pseudo_log_trans(sigma = 0.001))
p2 + labs(fill = "Population per square mile") + theme_map() +
    guides(fill = guide_legend(nrow = 1)) + theme(legend.position = "bottom")

```

## Refine your Plots

```{r}

library(ggrepel)
head(asasec)

p <- ggplot(data = subset(asasec, Year == 2014), mapping = aes(x = Members, y = Revenues, label = Sname))
p + geom_point() + geom_smooth()

p <- ggplot(data = subset(asasec, Year == 2014), mapping = aes(x = Members, y = Revenues, label = Sname))
p + geom_point(mapping = aes(color = Journal)) + geom_smooth(method = "lm")

p0 <- ggplot(data = subset(asasec, Year == 2014), mapping = aes(x = Members, y = Revenues, label = Sname))

p1 <- p0 + geom_point(mapping = aes(color = Journal)) + geom_smooth(method = "lm", se = FALSE, color = "gray80")

p2 <- p1 + geom_text_repel(data = subset(asasec, Year == 2014 & Revenues > 7000), size = 3)

p3 <- p2 + labs(x = "Membership", y = "Revenues", coloe = "Section has own Journal",
                title = "ASA Sections", subtitle = "2014 Calendar year.", caption = "Source: ASA annual report.")

p4 <- p3 + scale_y_continuous(labels = scales::dollar) +
    theme(legend.position = "bottom")
p4

```
```{r}

p <-ggplot(data = organdata, mapping = aes(x = roads, y = donors, color = world))

p + geom_point(size = 2) + scale_color_brewer(palette = "Set2") +
    theme(legend.position = "top")

p + geom_point(size = 2) + scale_color_brewer(palette = "Pastel2") +
    theme(legend.position = "top")

p + geom_point(size = 2) + scale_color_brewer(palette = "Dark2") +
    theme(legend.position = "top")

#demo('colors')     

cb_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

p4 + scale_color_manual(values = cb_palette)    # also possible: scale_color_hue()

library(dichromat)      # instead: library(colorblindr)

Default <- RColorBrewer::brewer.pal(5, "Set2")

types <- c("deutan", "protan", "tritan")
names(types) <- c("Deuteronopia", "Protanopia", "Tritanopia")

color_table <- types %>% purrr::map(~ dichromat(Default, .x)) %>%
    as_tibble() %>% add_column(Default, .before = TRUE)

color_table
color_comp(color_table)

```

```{r}

# Democrat Blue and Republican Red
party_colors <- c("#2E74C0", "#CB454A")

p0 <- ggplot(data = subset(county_data, flipped == "No"),
             mapping = aes(x = pop, y = black/100))

p1 <- p0 + geom_point(alpha = 0.15, color = "gray50") +
    scale_x_log10(labels = scales::comma)
p1

p2 <- p1 + geom_point(data = subset(county_data, flipped == "Yes"),
                      mapping = aes(x = pop, y = black/100, color = partywinner16)) +
    scale_color_manual(values = party_colors)
p2

p3 <- p2 + scale_y_continuous(labels = scales::percent) +
    labs(color = "County flipped to ...", x = "County Population(log scale)", y = "Percent Black Population",
         title = "Flipped counties, 2016", caption = "Counties in gray did not flip.")
p3

p4 <- p3 + geom_text_repel(data = subset(county_data, flipped == "Yes" & black > 25),
                           mapping = aes(x = pop, y = black/100, label = state), size = 2)
p4 + theme_minimal()  +
    theme(legend.position = "top")

theme_set(theme_bw())       # also: theme_classic(), theme_gray() and default theme_grey()
p4 + theme(legend.position = "top")

theme_set(theme_dark())
p4 + theme(legend.position = "top")

library(ggthemes)  # also: library(cowplot) and library(hrbrthemes)

theme_set(theme_economist())
p4 + theme(legend.position = "top")

theme_set(theme_wsj())
p4 + theme(legend.position = "top")

p4 + theme(plot.title = element_text(size = rel(0.6)),
           legend.title = element_text(size = rel(0.35)),
           plot.caption = element_text(size = rel(0.35)),
           legend.position = "top")

theme_set(theme_grey())
p4 + theme(legend.position = "top")

p4 + theme(legend.position = "top",
           plot.title = element_text(size = rel(2),
                                     lineheight = .5,
                                     family = "Times",
                                     face = "bold.italic",
                                     color = "orange"),
           axis.text.x = element_text(size = rel(1),
                                      family = "Courier",
                                      face = "bold",
                                      color = "purple"))

```
```{r, fig.width = 6, fig.height = 12}

yrs <- c(seq(1972, 1988, 4), 1993, seq(1996, 2016, 4))

mean_age <- gss_lon %>%
    filter(age %nin% NA && year %in% yrs) %>%
    group_by(year) %>%
    summarize(xbar = round(mean(age, na.rm = TRUE), 0))

mean_age$y <- 0.3
yr_labs <- data.frame(x = 85, y = 0.8, year = yrs)

p <- ggplot(data = subset(gss_lon, year %in% yrs),
            mapping = aes(x = age))

p1 <- p + geom_density(fill = "gray20", color = FALSE, alpha = 0.9, mapping = aes(y = ..scaled..)) +
    geom_vline(data = subset(mean_age, year %in% yrs),
               aes(xintercept = xbar), color = "white", size = 0.5) +
    geom_text(data = subset(mean_age, year %in% yrs),
              aes(x = xbar, y = y, label = xbar), nudge_x = 7.5, color = "white", size = 3.5, hjust = 1) +
    geom_text(data = subset(yr_labs, year %in% yrs),
              aes(x = x, y = y, label = year)) +
    facet_grid(year ~ ., switch = "y")
p1

library(hrbrthemes)
#library(myriad)

p1 + theme_ipsum(base_size = 10, plot_title_size = 10, # with library(myriad): theme_book, panel_spacing = unit(0.1, "lines")
                strip_text_size = 32) +
    theme(plot.title = element_text(size = 16),
          axis.text.x = element_text(size = 12),
          axis.title.y = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          strip.background = element_blank(),
          strip.text.y = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank()) +
    labs(x = "Age", y = NULL, title = "Age Distribution of\nGSS Respondents")

library(ggridges)

p <- ggplot(data = gss_lon,
            mapping = aes(x = age, y = factor(year, levels = rev(unique(year)), ordered = TRUE)))

p + geom_density_ridges(alpha = 0.6, fill = "lightblue", scale = 1.5) +
    scale_x_continuous(breaks = c(25, 50, 75)) +
    scale_y_discrete(expand = c(0.01, 0)) +
    labs(x = "Age", y = NULL, title = "Age Distribution of\nGSS Respondents") +
    theme_ridges() +
    theme(title = element_text(size = 16, face = "bold"))

```
```{r}

head(fredts)

fredts_m <- fredts %>% select(date, sp500_i, monbase_i) %>% gather(key = series, value = score, sp500_i:monbase_i)
head(fredts_m)

p <- ggplot(data = fredts_m, mapping = aes(x = date, y = score, group = series, color = series))

p1 <- p + geom_line() + theme(legend.position = "top") +
    labs(x = NULL, y = "Index", color = "Series")

p <- ggplot(data = fredts, mapping = aes(x = date, y = sp500_i - monbase_i))

p2 <- p + geom_line() +
    labs(x = "Date", y = "Difference")

cowplot::plot_grid(p1, p2, nrow = 2, rel_heights = c(0.75, 0.25), align = "v") # also: library(gridExtra) with grid.arrange()

```
```{r}

head(yahoo)

p <- ggplot(data = yahoo, mapping = aes(x = Employees, y = Revenue))

p + geom_path(color = "gray80") +
    geom_text(aes(color = Mayer, label = Year), size = 3, fontface = "bold") +
    theme(legend.position = "bottom") +
    labs(color = "Mayer is CEO", x = "Employees", y = "Revenue (Millions)",
         title = "Yahoo Employees vs Revenues, 2004-2014") +
    scale_y_continuous(labels = scales::dollar) +
    scale_x_continuous(labels = scales::comma)

p <- ggplot(data = yahoo, mapping = aes(x = Year, y = Revenue/Employees))

p + geom_vline(xintercept = 2012) +
    geom_line(color = "gray60", size = 2) +
    annotate("text", x = 2013, y = 0.44, label = "Mayer becomes CEO", size = 3) +
    labs(x = "Year", y = "Revenue/Employees", title = "Yahoo Revenue to Employee Ratio, 2004-2014")

```
```{r}

head(studebt)
studebt

p_xlab <- "Amount Owed, in thousands of Dollars"
p_title <- "Outstanding Student Loans"
p_subtitle <- "44 million borrowers owe a total of $1.3 trillion"
p_caption <- "Source: FRB NY"

f_labs <- c(`Borrowers` = "Percent of\nall Borrowers",
            `Balances` = "Percent of\nall Balances")

p <- ggplot(data = studebt, mapping = aes(x = Debt, y = pct/100, fill = type))

p + geom_bar(stat = "identity") +
    scale_fill_brewer(type = "qual", palette = "Dark2") +
    scale_y_continuous(labels = scales::percent) +
    guides(fill = FALSE) +
    theme(strip.text.x = element_text(face = "bold")) +
    labs(y = NULL, x = p_xlab, caption = p_caption, title = p_title, subtitle = p_subtitle) +
    facet_grid(~ type, labeller = as_labeller(f_labs)) +
    coord_flip()

library(viridis)

p <- ggplot(studebt, aes(y = pct/100, x = type, fill = Debtrc))

p + geom_bar(stat = "identity", color = "gray80") +
    scale_x_discrete(labels = as_labeller(f_labs)) +
    scale_y_continuous(labels = scales::percent) +
    scale_fill_viridis(discrete = TRUE) +
    guides(fill = guide_legend(reverse = TRUE,
                               title.position = "top", label.position = "bottom", keywidth = 3, nrow = 1)) +
    labs(x = NULL, y = NULL, fill = "Amount Owed, in thousands of dollars",
         caption = p_caption, title = p_title, subtitle = p_subtitle) +
    theme(legend.position = "top",
          axis.text.y = element_text(face = "bold", hjust = 1, size = 10),
          axis.ticks.length = unit(0, "cm"),
          panel.grid.major.y = element_blank()) +
    coord_flip()

```

## Appendix

```{r}

my_numbers <- c(1, 2, 3, 1, 3, 5, 25)
your_numbers <- c(5, 31, 71, 1, 3, 21, 6)

my_numbers[4]
my_numbers[7]
my_numbers[2:4]
my_numbers[c(2,4)]

my_tb <- tibble(mine = c(1, 4, 5, 8:11), yours = c(3, 20, 16, 34:31))
class(my_tb)
my_tb

my_tb[3, 1]         # row 3 col 1
my_tb[1, 2]         # row 1 col 2
my_tb[3, "mine"]    # row 3 col 1
my_tb[1, "yours"]   # row 1 col 2
my_tb[, "mine"]     # all rows col 1
my_tb[4, ]          # row 4 all cols

my_tb$mine

out <- lm(mine ~ yours, data = my_tb)
out$coefficients
out$call
out$qr$rank         # nested

my_tb$ours <- my_tb$mine + my_tb$yours
my_tb

edu

edu_t <- gather(data = edu, key = school, value = freq, elem4:coll4)

head(edu_t)
tail(edu_t)

```
```{r}
 
head(bad_date)

p <- ggplot(data = bad_date, aes(x = date, y = N))
p + geom_line()

bad_date2 <- rbind(bad_date, bad_date)

p <- ggplot(data = bad_date2, aes(x = date, y = N))
p + geom_line()

#install.packages('lubridate')
library('lubridate')

bad_date3 <- bad_date
bad_date3$date <- mdy(bad_date$date)
head(bad_date3)

p <- ggplot(data = bad_date3, aes(x = date, y = N))
p + geom_line()

url <- "https://cdn.rawgit.com/kjhealy/viz-organdata/master/organdonation.csv"
bad_year <- read_csv(url)
bad_year %>% select(1:3) %>% sample_n(10)

p <- ggplot(data = bad_year, aes(x = year, y = donors))
p + geom_point()

bad_year$year <- int_to_year(bad_year$year)
bad_year %>% select(1:3)

p <- ggplot(data = bad_year, aes(x = year, y = donors))
p + geom_point()

```
```{r}

add_xy <- function(x, y) {
    x + y
}

add_xy(x = 1, y = 7)
add_xy(5, 2)

plot_section <- function(section = "Culture", x = "Year", y = "Members", data = asasec, smooth = FALSE) {
    
    require(ggplot2)
    require(splines)
    
    p <- ggplot(subset(data, Sname == section), mapping = aes_string(x = x, y = y))     # aes_string() instead of aes()
    
    if(smooth == TRUE) {
        
        p0 <- p + geom_smooth(color = "#999999", size = 1.2, method = "lm", formula = y ~ ns(x, 3)) +
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            labs(title = section)
        
    } else {
        
        p0 <- p + geom_line(color = "#E69F00", size = 1.2) +
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            labs(title = section)
        
    }
    
    print(p0)
    
}

plot_section("Rationality")
plot_section("Sexualities", smooth = TRUE)

plot_section <- function(section = "Culture", x = "Year", y = "Members", data = asasec, smooth = FALSE, ...) {
    
    require(ggplot2)
    require(splines)
    
    p <- ggplot(subset(data, Sname == section), mapping = aes_string(x = x, y = y))     # aes_string() instead of aes()
    
    if(smooth == TRUE) {
        
        p0 <- p + geom_smooth(color = "#999999", size = 1.2, ...) +
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            labs(title = section)
        
    } else {
        
        p0 <- p + geom_line(color = "#E69F00", size = 1.2) +
            scale_x_continuous(breaks = c(seq(2005, 2015, 4))) +
            labs(title = section)
        
    }
    
    print(p0)
    
}

plot_section("Comm/Urban", smooth = TRUE, method = "loess")
plot_section("Children", smooth = TRUE, method = "lm", formula = y ~ ns(x, 2))

```